---
title: "RAROC-Analysis"
output: html_document
date: "2025-04-15"
---

#Loading packages
```{r}
library(dplyr)
library(reshape)
library(ggplot2)
library(nlme)
library(Hmisc)
library(tidyverse)
library(stats)
library(scales)
library(FinCal)
library(waterfalls)
library(cowplot)
library(grid)
library(patchwork)
library(gridGraphics)
```

#Basic setup
```{r}
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
options(tidyverse.quiet = TRUE)

ITFcolours<- c(
  "#1f77b4",  # strong blue
  "#d62728",  # bold red
  "#5fa00a",  # strong green
  "#ff7f0e",  # vivid orange
  "#9867dd",  # medium purple
  "#aaa000",  # lime yellow
  "#081ed3",  # turquoise
  "#390b79",  # deep indigo
  "#af1493",  # deep pink
  "#ff1147",  # tomato red
  "#2A6B1A",  # forest green
  "#ffcc00",  # bright yellow
  "#19998f",  # teal green
  "#e377c2",  # vibrant pink
  "#9f4500"   # orange-red
)

# Working directory: anchored on this Rmd's folder (portable across machines)
get_script_dir <- function() {
  ki <- try(knitr::current_input(dir = TRUE), silent = TRUE)
  if (!inherits(ki, "try-error") && !is.null(ki) && nzchar(ki)) return(dirname(ki))
  if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
    p <- rstudioapi::getSourceEditorContext()$path
    if (nzchar(p)) return(dirname(p))
  }
  getwd()
}
script_dir <- get_script_dir()
```

#Preparing the input data
```{r}
input_dir   <- file.path(script_dir, "Inputs")
raroc_input <- file.path(script_dir, "Inputs", "RAROC")
output_dir  <- script_dir
```

```{r}
#Other inputs from https://theicct.org/wp-content/uploads/2021/11/tco-bets-europe-1-nov21.pdf

#General for NPV:
ownership_time <- 5 #years
discount_rate <- 0.095
discount_rate_public <- 0.045
interest_rate <- 0.06
upfront_cost <- 0.2 #Variable added for sensitivity analysis
residual_value <- 0.3 #Diesel
subsidy <- 0 #Subsidy

#Fixed costs:
registration_fee<- 800 #[Eur] // Value for France
DT_price <- data.frame( #[Eur]
  year = c(2024, 2030),
  price = c(134100, 137300)) 

BET_price <- read.csv(file.path(input_dir, "BET-Price-IEA_adjusted.csv")) %>% #Source: https://iea.blob.core.windows.net/assets/0aa4762f-c1cb-4495-987a-25945d6de5e8/GlobalEVOutlook2025.pdf adjusted with projections from https://theicct.org/wp-content/uploads/2021/11/tco-bets-europe-1-nov21.pdf
  dplyr::rename(year = Year, price = BET.Retail.Price..Eur.)
  
#Variable costs or parameters:
ownership_tax <- 950 #[Eur/year] // Value for France
road_use_charge <- 0 #[Eur/year] // Value for France
road_tolls <- 0.32 #[Eur/km] // Value for France
maintenance_costs <- data.frame( #[Eur/100km]
  fuel = c("diesel", "BET"),
  Maintenance_costs = c(18.5, 13.24)
)

diesel_consumption <- data.frame( #[l/100km]
  year = c(2024, 2030),
  fuel_consumption = c(27.7, 27.15) #Source: IICT adjusted with IEA trends 
)

energy_consumption <- data.frame( #[kWh/km]
  year = c(2020, 2030),
  energy_consumption = c(1.38, 0.99)
  )

diesel_price <- data.frame( #[Eur/l]
  year= c(2020,2030),
  price = c(0.905, 0.905)
)

electricity_price <- data.frame( #[Eur/kWh]
  year = c(2020, 2030),
  price = c(0.369, 0.127)
)

VKT_year <- read.csv(file.path(input_dir, "VKT_perYear.csv")) %>% #Source: https://theicct.org/wp-content/uploads/2021/11/tco-bets-europe-1-nov21.pdf
  dplyr::rename(service_year = Service.year, VKT = VKT..km.)
```


#Interpolating data between years

```{r}
years_to_interpolate = 2020:2035

DT_price <- approx(x=DT_price$year, y=DT_price$price, xout = 2024:2030, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, price = y)
 
diesel_consumption <- approx(x=diesel_consumption$year, y=diesel_consumption$fuel_consumption, xout = 2024:2035, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, fuel_consumption = y)

energy_consumption <- approx(x=energy_consumption$year, y=energy_consumption$energy_consumption, xout = years_to_interpolate, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, energy_consumption = y)

diesel_price <- approx(x=diesel_price$year, y=diesel_price$price, xout = years_to_interpolate, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, price = y)

electricity_price <- approx(x=electricity_price$year, y=electricity_price$price, xout = years_to_interpolate, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, price = y)
```

## -------- GET incentive to reduce TCO -----

```{r}
get_TCO_detailed <- function(Purchase_year, interest, Fuel, upfront_cost, residual_value, subsidy_type, subsidy, road_tolls_exemp){
    
  if (Fuel == "diesel"){
    vehicle_price <- DT_price$price[DT_price$year==Purchase_year]
  }else if (Fuel == "BET"){
    vehicle_price <- BET_price$price[BET_price$year==Purchase_year]
  }
  
  #SUBSIDY
  if(subsidy_type == "fixed_percentage") {
    subsidy_value <- subsidy*vehicle_price
  } else if (subsidy_type == "percentage_difference") {
    vehicle_price_diesel <- DT_price$price[DT_price$year==Purchase_year]
    subsidy_value <- (vehicle_price - vehicle_price_diesel)*subsidy
  } else if (subsidy_type == "fixed_amount"){
    vehicle_price_diesel <- DT_price$price[DT_price$year==Purchase_year]
    if (subsidy <= (vehicle_price - vehicle_price_diesel)){
      subsidy_value <- subsidy
    } else {
      subsidy_value <- (vehicle_price - vehicle_price_diesel)
    }
  }
  
  total_loan <- (vehicle_price-subsidy_value)*(1-upfront_cost)
  annual_payment <- pmt(
    r=interest/12, n=ownership_time*12, pv=-total_loan, fv=0)*12 # I removed fv= vehicle_price*residual_value
  
  cash_flow_detailed <- data.frame()
  
  for (year_ow in 0:ownership_time) {
    
    #Checking toll exemption
    if(Purchase_year + year_ow <= road_tolls_exemp){
      toll_exemp <- 0 #Exempted
    }else {
      toll_exemp <- 1 #No_exempted
    }
    
    if (year_ow == 0) {
      vehicle_purchase <- (vehicle_price-subsidy_value)*upfront_cost
      taxes <- registration_fee
      maintenance <- 0
      tolls <- 0
      fuel_consumption <- 0
      
    } else {
      VKT <- VKT_year$VKT[VKT_year$service_year == year_ow]
      
      vehicle_purchase <- annual_payment
      taxes <- ownership_tax + road_use_charge
      maintenance <- maintenance_costs$Maintenance_costs[maintenance_costs$fuel == Fuel]*VKT/100
      tolls <- road_tolls*VKT*toll_exemp
      
      if (Fuel == "diesel") {
        fuel_consumption <- diesel_consumption$fuel_consumption[diesel_consumption$year == Purchase_year] *
          VKT * diesel_price$price[diesel_price$year == (Purchase_year + year_ow - 1)] / 100
      } else if (Fuel == "BET") {
        fuel_consumption <- energy_consumption$energy_consumption[energy_consumption$year == Purchase_year] *
          VKT * electricity_price$price[electricity_price$year == (Purchase_year + year_ow - 1)]
      }
      
    }
    
    if (year_ow == ownership_time) {
      vehicle_purchase <- vehicle_purchase-vehicle_price*residual_value
    } else {
      vehicle_purchase <- vehicle_purchase
    }
    
    cash_flow_detailed <- rbind(
      cash_flow_detailed,
      data.frame(
        year = year_ow,
        vehicle_purchase = vehicle_purchase,
        taxes = taxes,
        maintenance = maintenance,
        tolls = tolls,
        fuel_consumption = fuel_consumption
        #vehicle_resale = vehicle_resale
      )
    )
    
  }
  
  NPV_detailed = data.frame(
    vehicle_purchase = npv(r = discount_rate, cf = cash_flow_detailed$vehicle_purchase),
    taxes = npv(r = discount_rate, cf = cash_flow_detailed$taxes),
    maintenance = npv(r = discount_rate, cf = cash_flow_detailed$maintenance),
    tolls = npv(r = discount_rate, cf = cash_flow_detailed$tolls),
    fuel_consumption = npv(r = discount_rate, cf = cash_flow_detailed$fuel_consumption)
    #vehicle_resale = npv(r = discount_rate, cf = cash_flow_detailed$vehicle_resale)
  ) %>%
    mutate(NPV = vehicle_purchase + taxes + maintenance + tolls + fuel_consumption) # + vehicle_resale)
  
  
  return(NPV_detailed)
  
}
```


```{r}
get_NPV_fin_instrument <- function(Purchase_year, subsidy_type, subsidy, interest_high, interest_low, residual_value, road_tolls_exemp, second_hand_value) {
  
  vehicle_price <- BET_price$price[BET_price$year==Purchase_year] #Calculating the vehicle price
  vehicle_price_diesel <- DT_price$price[DT_price$year==Purchase_year] #Calculating the price of diesel truck
  
  
  #SUBSIDY
  if(subsidy_type == "fixed_percentage") {
    subsidy_value <- subsidy*vehicle_price
  } else if (subsidy_type == "percentage_difference") {
    subsidy_value <- (vehicle_price - vehicle_price_diesel)*subsidy
  } else if (subsidy_type == "fixed_amount"){
    if (subsidy <= (vehicle_price - vehicle_price_diesel)){
      subsidy_value <- subsidy
    } else {
      subsidy_value <- (vehicle_price - vehicle_price_diesel)
    }
  }

  total_loan <- (vehicle_price-subsidy_value)*(1-upfront_cost) #Estimating the total loan
  annual_payment_high <- pmt( #Annual payment with high interest
    r=interest_high/12, n=ownership_time*12, pv=-total_loan, fv = 0)*12 ## Here, in pv I am making interests independent to the subsidy because I want to evaluate the financial instruments individually. 
  
  annual_payment_low <- pmt( #Annual payment with low interest
    r=interest_low/12, n=ownership_time*12, pv=-total_loan, fv =0)*12 ## Here, in pv I am making interests independent to the subsidy because I want to evaluate the financial instruments individually. 
  
  financial_instrument <- data.frame() 
   
  for (year_ow in 0:ownership_time) { #Running the calculations for the ownership period
    
    #Checking toll exemption
    if(Purchase_year + year_ow <= road_tolls_exemp){
      toll_exemp <- 0 #Exempted
    }else {
      toll_exemp <- 1 #No_exempted
    }
    
    if (year_ow == 0) { 
      
      remaining_balance_high <- (vehicle_price-subsidy_value)*upfront_cost #Here, I am making interests independent to the subsidy because I want to evaluate the financial instruments individually. 
      remaining_balance_low <- (vehicle_price-subsidy_value)*upfront_cost #Here, I am making interests independent to the subsidy because I want to evaluate the financial instruments individually.
      purchase_subsidy = subsidy_value
      tolls_subsidy <- 0
      interest_subsidy <- 0
      residual_value_subsidy <- 0
      
    } else {
      VKT <- VKT_year$VKT[VKT_year$service_year == year_ow]
      
      purchase_subsidy <- 0
      tolls_subsidy <- road_tolls*VKT*(1-toll_exemp)
      
      interest_high_paid <- remaining_balance_high * interest_high #Interest paid with high interest
      interest_low_paid <- remaining_balance_low * interest_low #Interest paid with  low interest
      
      interest_subsidy <- interest_high_paid - interest_low_paid #Subsidy on interests
      
      remaining_balance_high <- remaining_balance_high - (annual_payment_high - interest_high_paid) #Remaining debt balance to recalculate interests
      remaining_balance_low <- remaining_balance_low - (annual_payment_low - interest_low_paid)
      
      if (year_ow == ownership_time) {
        residual_value_subsidy <- max(0,vehicle_price*(residual_value - second_hand_value)) #Only considered in the last year and if no zero
      } else {
        residual_value_subsidy <- 0
      }

    }
    
    financial_instrument <- rbind( 
      financial_instrument,
      data.frame(
        year = year_ow,
        purchase_subsidy = purchase_subsidy,
        tolls_subsidy = tolls_subsidy,
        interest_subsidy = interest_subsidy,
        residual_value_subsidy = residual_value_subsidy
      )
    )
    
  }
  
  NPV_financial_instrument = data.frame( #Calculate NPV
    purchase_subsidy = npv(r = discount_rate_public, cf = financial_instrument$purchase_subsidy),
    tolls_subsidy = npv(r = discount_rate_public, cf = financial_instrument$tolls_subsidy),
    interest_subsidy = npv(r = discount_rate_public, cf = financial_instrument$interest_subsidy),
    residual_value_subsidy = npv(r = discount_rate_public, cf = financial_instrument$residual_value_subsidy)
  ) 
  
  return(NPV_financial_instrument)
}
```

##--------------- RAROC AND RoE ------------

```{r}
PD <- readRDS(file.path(raroc_input, "PD_WB.rds"))

recovery_rate <- readRDS(file.path(raroc_input, "RecoveryRate_WB.rds"))


CB <- readRDS(file.path(raroc_input, "CentralBank_WB.rds")) %>%
  mutate(Value = Value/100) %>%
  filter(!is.na(Region.wb))

OpCosts <- readRDS(file.path(raroc_input, "OperationalCost_WB.rds"))
```

#FUNCTION TO CALCULATE THE INTEREST NEEDED TO REACH CERTAIN LEVEL OF RETURN ON EQUITY

```{r}
## Followed methodology: https://roopya.money/introduction-to-loan-pricing-and-raroc-risk-adjusted-return-on-capital/
###OUTPUT: INTEREST RATE
get_interest_raroc <- function(Purchase_year, Fuel, upfront_cost, subsidy, residual_value, PD, region, company){
  
  if (company == "SME") {
    data_RR <- recovery_rate %>% 
      select(Region, RR_industrial) %>%
      dplyr::rename(RR = "RR_industrial")
  } else if (company == "industry"){
    data_RR <- recovery_rate %>% 
      select(Region, RR_SMEs) %>%
      dplyr::rename(RR = "RR_SMEs")
  } else if (company == "Market"){
    data_RR <- recovery_rate %>% 
      select(Region, RR)
  }
  
  if (Fuel == "Diesel") {
    vehicle_price <- DT_price$price[DT_price$year == Purchase_year]
    
    RecoveryRate <- data_RR$RR[data_RR$Region == region]
    
  } else if (Fuel == "BET") {
    vehicle_price <- BET_price$price[BET_price$year == Purchase_year]
    RecoveryRate <- residual_value
  }
  
  total_loan <- (1 - upfront_cost - subsidy) * vehicle_price 

  
  LGD <- 1 - RecoveryRate #Source: https://www.investopedia.com/terms/l/lossgivendefault.asp
  EL <- total_loan*PD*LGD #Expected loss
  
  Economic_capital_requirement <- 0.08
  
  Cost_capital <- CB$Value[CB$Region.wb == region]

  Economic_capital = Economic_capital_requirement*total_loan
  
  raroc <- 0.1
  
  Op_expenses <- OpCosts$Value[OpCosts$Region.wb == region]

  NII <- (raroc*Economic_capital + EL)/(1-Op_expenses)
  
  interest_rate_output <- (NII + Cost_capital*Economic_capital)/total_loan
  
  return(interest_rate_output)
}
```

# Function to calculate the incentive required to reach a TCO reduction goal

```{r}
find_incentive <- function(TCO_reduction_goal, year, incentive_type, guess_value) {

  residual_value <- RV_market$RV[RV_market$year == year]
  road_tolls_exemp <- 0
  subsidy <- 0
  
  baseline_TCO <- BET_tco$NPV[BET_tco$year == year]
  target_TCO <- baseline_TCO * (1-TCO_reduction_goal)
  
  # Update the parameter based on the input
  if (incentive_type == "subsidy") {
    subsidy <- guess_value
    PD_BET_find <- PD_BET_before$PD[PD_BET_before$year==year]
      
  } else if (incentive_type == "residual_value") {
    residual_value <- guess_value
    PD_BET_find <- PD_BET_after$PD[PD_BET_after$year==year]
    
  } else if (incentive_type == "tolls") {
    road_tolls_exemp <- guess_value
  } else if (incentive_type == "combined") { #Combined means RVG + subsidy
    residual_value <- 0.25
    subsidy <- guess_value
    PD_BET_find <- PD_BET_after$PD[PD_BET_after$year==year]

  }
  
  interest_new <- get_interest_raroc(Purchase_year = year, 
                                     Fuel="BET", 
                                     upfront_cost = upfront_cost, 
                                     subsidy = subsidy, 
                                     residual_value = residual_value, 
                                     PD = PD_BET_find, 
                                     region = "Europe", 
                                     company = "industry")
  
  
  #Estimating the new TCO
  TCO_estimated <- get_TCO_detailed(
    Purchase_year = year,
    interest = interest_new,
    # interest = interest_rate,
    Fuel = "BET",
    upfront_cost = upfront_cost,
    residual_value = residual_value,
    subsidy_type = "fixed_percentage", 
    subsidy = subsidy,
    road_tolls_exemp = road_tolls_exemp
  )
  
  return(TCO_estimated$NPV-target_TCO)
}
```

```{r}
PD_BET_before<- data.frame(
  year = 2024:2035) %>%
  mutate(PD =  PD$PD_industrial[PD$Region == "Europe"] /(1 + exp(2*(year - 2030))) + PD$PD_industrial[PD$Region == "Europe"])

ggplot()+
  geom_line(data=PD_BET_before, aes(x=year, y=PD*100, color = "PD"), size = 1.2) +
  scale_color_manual(values = ITFcolours) +
  scale_fill_manual(values = ITFcolours) +
  scale_y_continuous(expand = c(0,0),
                     limits=c(0,6)) +
  scale_x_continuous(expand = c(0,0),
                     labels = number_format(accuracy = 1, big.mark = "")) +
  labs(title="Probability of Default over time", 
       y="PD (%)", 
       x="Year") +
  theme_classic() +
  theme(legend.position = "none")
```


```{r}

PD_BET_after <- PD_BET_before

ggplot()+
  geom_line(data=PD_BET_after, aes(x=year, y=PD*100, color = "PD"), size = 1.2) +
  scale_color_manual(values = ITFcolours) +
  scale_fill_manual(values = ITFcolours) +
  scale_y_continuous(expand = c(0,0),
                     limits=c(0,6)) +
  scale_x_continuous(expand = c(0,0),
                     labels = number_format(accuracy = 1, big.mark = "")) +
  labs(title="Probability of Default over time",
       y="PD (%)",
       x="Year") +
  theme_classic() +
  theme(legend.position = "none")
```


# Estimating market RV with no incentive
```{r}
RV_market <- data.frame(
  year = 2024:2040) %>%
  mutate(RV = 0.20 /(1 + exp(-1*(year - 2030))))

ggplot()+
  geom_line(data=RV_market, aes(x=year, y=RV*100, color = "RV"), size = 1.2) +
  scale_color_manual(values = ITFcolours) +
  scale_fill_manual(values = ITFcolours) +
  scale_y_continuous(expand = c(0,0),
                     limits=c(0,25)) +
  scale_x_continuous(expand = c(0,0),
                     labels = number_format(accuracy = 1, big.mark = "")) +
  labs(title="Market RV over time with no incentive", 
       y="RV (%)", 
       x="Year") +
  theme_classic() +
  theme(legend.position = "none")
```

#Calculating BET/diesel TCO ratio updating IRs:
```{r}
diesel_tco <- data.frame()
BET_tco <- data.frame()

tco_BET_diesel <- data.frame()

for (year_purchase in 2024:2030){
  
  #Diesel TCO
  result_tco_diesel <- get_TCO_detailed(Purchase_year = year_purchase, Fuel = "diesel", interest= interest_rate, upfront_cost = upfront_cost, residual_value = residual_value, subsidy_type = "fixed_percentage", subsidy = subsidy, road_tolls_exemp= 0) #Diesel TCO
  
  #New interest rates for BET with no incentive:
  interest_new <- get_interest_raroc(Purchase_year = year_purchase, Fuel="BET", upfront_cost = upfront_cost, subsidy = 0, residual_value = RV_market$RV[RV_market$year==year_purchase], PD = PD_BET_before$PD[PD_BET_before$year==year_purchase], region = "Europe", company = "industry") 
  
  #BET TCO
  result_tco_BET <- get_TCO_detailed(
    Purchase_year = year_purchase, 
    interest = interest_new,
    # interest = interest_rate,
    Fuel = "BET",
    upfront_cost = upfront_cost,
    residual_value = RV_market$RV[RV_market$year==year_purchase],
    subsidy_type = "fixed_percentage",
    subsidy = 0,
    road_tolls_exemp = 0
  )
  
  tco_BET_diesel <- rbind(tco_BET_diesel, data.frame(year = year_purchase, (result_tco_BET/result_tco_diesel$NPV)))
  
  diesel_tco <- rbind(diesel_tco, data.frame(year = year_purchase, result_tco_diesel))

  BET_tco <- rbind(BET_tco, data.frame(year = year_purchase, result_tco_BET))
  
  }
```

```{r}
#Plotting TCO difference:
ggplot(tco_BET_diesel, aes(x = year, y = NPV)) +
  geom_line(aes(color = "BET/Diesel Ratio"), size = 1.2) + 
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", size = 1) +
  scale_color_manual(values=ITFcolours) +
  scale_y_continuous(#expand = c(0,0),
                     labels = comma) +
  scale_x_continuous(expand = c(0,0),
                     labels = number_format(accuracy = 1, big.mark = "")) +
  labs(title="Ratio of BET/Diesel TCO 2024-2030", y="TCO ratio BET/Diesel (-)", 
       x="Year of purchase",
       color = "")+
  theme_classic()
```

#Calculating the incentive to reach TCO parity

```{r}
subsidy_results <- data.frame()

max_subsidy <- 0.5 #Maximum subsidy possible

for (year in 2024:2030) {
  
  #Calculating the new IR
  IR_new <- get_interest_raroc(Purchase_year = year, 
                               Fuel="BET", 
                               upfront_cost = upfront_cost, 
                               subsidy = max_subsidy, 
                               residual_value = RV_market$RV[RV_market$year==year], 
                               PD = PD_BET_before$PD[PD_BET_before$year==year], 
                               region = "Europe", 
                               company = "industry")
  
    subsidy_solution <- max_subsidy 
    
    #Calculating the actual TCO with the max subsidy
    TCO_BET <- get_TCO_detailed(
      Purchase_year = year, 
      interest = IR_new,
      # interest = interest_rate,
      Fuel = "BET",
      upfront_cost = upfront_cost,
      residual_value = RV_market$RV[RV_market$year==year],
      subsidy_type = "fixed_percentage",
      subsidy = subsidy_solution,
      road_tolls_exemp = 0
    )$NPV
    
    if (TCO_BET <= diesel_tco$NPV[diesel_tco$year == year]){
      TCO_parity <- "Yes"
    } else {
      TCO_parity <- "No"
    }
    
    #Calculating the actual TCO reduction
    TCO_red <- (BET_tco$NPV[BET_tco$year == year]/TCO_BET)-1
  
  NPV_gov = get_NPV_fin_instrument(Purchase_year=year, 
                                   subsidy_type = "fixed_percentage", 
                                   subsidy = subsidy_solution, 
                                   interest_high = IR_new, 
                                   interest_low = IR_new, 
                                   residual_value = 0, 
                                   road_tolls_exemp=0, 
                                   second_hand_value=RV_market$RV[RV_market$year==year])
  
  subsidy_results <- rbind(subsidy_results,
                           data.frame( year = year, 
                                       subsidy_req = subsidy_solution, 
                                       NPV_gov = NPV_gov$purchase_subsidy,
                                       TCO_parity = TCO_parity,
                                       TCO_red = TCO_red))
}

```

# Residual value guarantee
```{r}
rvg_results <- data.frame()

max_rvg <- 0.25 #Maximum RVG

for (year in 2024:2030) {
  
  second_hand_value <- RV_market$RV[RV_market$year==year]
  
  #Calculating the new IR
  IR_new <- get_interest_raroc(Purchase_year = year, 
                               Fuel="BET", 
                               upfront_cost = upfront_cost, 
                               subsidy = 0, #No subsidy
                               residual_value = max_rvg,
                               PD = PD_BET_after$PD[PD_BET_after$year==year], 
                               region = "Europe", 
                               company = "industry")
  
  rvg_solution <- max_rvg #Max RVG possible

  #Calculating the actual TCO with the max subsidy
  TCO_BET <- get_TCO_detailed(
    Purchase_year = year, 
    interest = IR_new,
    # interest = interest_rate,
    Fuel = "BET",
    upfront_cost = upfront_cost,
    residual_value = rvg_solution,
    subsidy_type = "fixed_percentage",
    subsidy = 0,
    road_tolls_exemp = 0
  )$NPV
  
  if (TCO_BET <= diesel_tco$NPV[diesel_tco$year == year]){
    TCO_parity <- "Yes"
  } else {
    TCO_parity <- "No"
  }
  
  #Calculating the actual TCO reduction
  TCO_red <- (BET_tco$NPV[BET_tco$year == year]/TCO_BET) -1
  
  NPV_gov = get_NPV_fin_instrument(Purchase_year=year, 
                                   subsidy_type = "fixed_percentage", 
                                   subsidy = 0, 
                                   interest_high = IR_new, 
                                   interest_low = IR_new, 
                                   residual_value = rvg_solution, 
                                   road_tolls_exemp=0, 
                                   second_hand_value=RV_market$RV[RV_market$year==year+5]) # Cost for governments with a RVG at the 5 year)
  
  rvg_results <- rbind(rvg_results,
                       data.frame( year = year, 
                                   rvg_req = rvg_solution,
                                   second_hand_value =RV_market$RV[RV_market$year==year+5], # Cost for governments with a RVG at the 5 year,
                                   NPV_gov = NPV_gov$residual_value_subsidy,
                                   TCO_parity = TCO_parity,
                                   TCO_red = TCO_red))
  
}

```

# Combined measures: RVG+Subsidy
```{r}
combined_results <- data.frame()
max_subsidy <- 0.335 #Maximum subsidy possible
max_rvg <- 0.25

for (year in 2024:2030) { #Only consider the years where the RVG was higher to the maximum RVG set in 25%
  
  second_hand_value <- RV_market$RV[RV_market$year==year]

  #Calculating the new IR
  IR_new <- get_interest_raroc(Purchase_year = year, 
                               Fuel="BET", 
                               upfront_cost = upfront_cost, 
                               subsidy = max_subsidy, #No subsidy
                               residual_value = max_rvg, #Max between Market RV a 25%
                               PD = PD_BET_after$PD[PD_BET_after$year==year], 
                               region = "Europe", 
                               company = "industry")
  
  combined_solution <- max_subsidy #Max subsidy possible

  #Calculating the actual TCO with the max subsidy
  TCO_BET <- get_TCO_detailed(
    Purchase_year = year, 
    interest = IR_new,
    Fuel = "BET",
    upfront_cost = upfront_cost,
    residual_value = max_rvg,
    subsidy_type = "fixed_percentage",
    subsidy = max_subsidy,
    road_tolls_exemp = 0
  )$NPV
  
  if (TCO_BET <= diesel_tco$NPV[diesel_tco$year == year]){
    TCO_parity <- "Yes"
  } else {
    TCO_parity <- "No"
  }
  
  #Calculating the actual TCO reduction
  TCO_red <- (BET_tco$NPV[BET_tco$year == year]/TCO_BET) -1

  NPV_gov = get_NPV_fin_instrument(Purchase_year=year, 
                                   subsidy_type = "fixed_percentage", 
                                   subsidy = combined_solution, 
                                   interest_high = IR_new, 
                                   interest_low = IR_new, 
                                   residual_value = max_rvg, 
                                   road_tolls_exemp=0, 
                                   second_hand_value=RV_market$RV[RV_market$year==year+5]) # Cost for governments with a RVG at the 5 year
  
  combined_results <- rbind(combined_results,
                            data.frame( year = year, 
                                        rvg_req = max_rvg,
                                        subsidy_req = combined_solution,
                                        second_hand_value =RV_market$RV[RV_market$year==year+5], # Cost for governments with a RVG at the 5 year,
                                        NPV_gov = NPV_gov$residual_value_subsidy + NPV_gov$purchase_subsidy,
                                        TCO_parity = TCO_parity,
                                        TCO_red = TCO_red))
}

```

```{r}
subsidy_results_gg <- subsidy_results %>%
  arrange(year) %>%
  mutate(year_next = lead(year),
         NPV_next = lead(NPV_gov),
         TCO_parity_next = lead(TCO_parity)) %>%
  filter(!is.na(year_next))

rvg_results_gg <- rvg_results %>%
  arrange(year) %>%
  mutate(year_next = lead(year),
         NPV_next = lead(NPV_gov),
         TCO_parity_next = lead(TCO_parity)) %>%
  filter(!is.na(year_next))

combined_results_gg <- combined_results %>%
  arrange(year) %>%
  mutate(year_next = lead(year),
         NPV_next = lead(NPV_gov),
         TCO_parity_next = lead(TCO_parity)) %>%
  filter(!is.na(year_next))

```

```{r}
## ESTIMATING WHEN TCO PARITY IS REACHED:

TCO_parity_subsidy <- subsidy_results %>%
  filter(TCO_parity == "Yes") %>%
  slice_min(year, n = 1) %>%
  select(year, NPV_gov) %>%
  mutate(incentive = "Subsidy")

TCO_parity_rvg <- rvg_results %>%
  filter(TCO_parity == "Yes") %>%
  slice_min(year, n = 1) %>%
  select(year, NPV_gov) %>%
  mutate(incentive = "RVG")

TCO_parity_combined <- combined_results %>%
  filter(TCO_parity == "Yes") %>%
  slice_min(year, n = 1) %>%
  select(year, NPV_gov) %>%
  mutate(incentive = "RVG + Subsidy")

TCO_parity_gg <- bind_rows(TCO_parity_subsidy,
                           TCO_parity_rvg,
                           TCO_parity_combined)
```



```{r}
Gov_expenditure_combined_measures <- ggplot() +
  geom_segment(data=subsidy_results_gg,
               aes(x = year, xend = year_next,
                   y = NPV_gov, yend = NPV_next, 
                   color = factor("Subsidy (50%)", 
                                  levels = c("Subsidy (50%)", 
                                             "Subsidy (30%) + RVG (25%)", 
                                             "RVG (25%)")), 
                   linetype = TCO_parity), 
               size = 1.2) +
  
  geom_segment(data=combined_results_gg,
               aes(x = year, xend = year_next,
                   y = NPV_gov, yend = NPV_next,
                   color = factor("Subsidy (30%) + RVG (25%)", 
                                  levels = c("Subsidy (50%)", 
                                             "Subsidy (30%) + RVG (25%)", 
                                             "RVG (25%)")),
                   linetype = TCO_parity),
               size = 1.2) +
  
  geom_segment(data=rvg_results_gg,
               aes(x = year, xend = year_next,
                   y = NPV_gov, yend = NPV_next,
                   color = factor("RVG (25%)", 
                                  levels = c("Subsidy (50%)", 
                                             "Subsidy (30%) + RVG (25%)", 
                                             "RVG (25%)")),
                   linetype = TCO_parity),
               size = 1.2) +
  
  geom_point(data= TCO_parity_gg,
             aes(x = year,
                 y = NPV_gov, shape = "Onset of TCO parity"),
             color = "black",
             size = 3.5)+
  
  scale_color_manual(values = ITFcolours) +
  scale_fill_manual(values = ITFcolours) +
  scale_linetype_manual(values = c("Yes" = "solid",
                                   "No" = "dotted")) +
  scale_y_continuous(labels = function(x) format(x, big.mark = " "),
                     expand = c(0,0),
                     limits = c(0,200000))+
  scale_x_continuous(expand = c(0,0),
                     limits = c(2024, 2030.1)) +
  labs(
       y="Government expenditure (€/BET)", 
       x="Year",
       color = "Financial incentive \n(% of initial purchase price)",
       linetype = "TCO parity",
       shape = "") +
    theme_classic()#+

Gov_expenditure_combined_measures
```
# SENSITIVITY ANALYSIS: RV_market ±5pp, PD_BET_after ±2pp 

```{r}

rv_deltas <- c(-0.05, 0.05)

max_subsidy_s <- 0.5
max_rvg_s <- 0.25
max_subsidy_combined_s <- 0.30
max_rvg_combined_s <- 0.25

#1. Variations in RV_market (PD at base)
rv_sub <- data.frame()
rv_rvg <- data.frame()
rv_com <- data.frame()

for (rv_d in rv_deltas) {

  # RV_adj <- RV_market %>% mutate(RV = pmax(0, RV + rv_d))
  RV_adj <- RV_market %>% mutate(RV = rv_d)

  for (year in 2024:2030) {

    #  Subsidy 
    IR_s <- get_interest_raroc(
      Purchase_year = year, Fuel = "BET",
      upfront_cost = upfront_cost, subsidy = max_subsidy_s,
      residual_value = RV_adj$RV[RV_adj$year == year],
      PD = PD_BET_before$PD[PD_BET_before$year == year],
      region = "Europe", company = "industry")

    NPV_s <- get_NPV_fin_instrument(
      Purchase_year = year, subsidy_type = "fixed_percentage",
      subsidy = max_subsidy_s, interest_high = IR_s, interest_low = IR_s,
      residual_value = 0, road_tolls_exemp = 0,
      second_hand_value = RV_adj$RV[RV_adj$year == year])

    rv_sub <- rbind(rv_sub, data.frame(year = year, rv_d = rv_d,
                                       NPV_gov = NPV_s$purchase_subsidy))

    # RVG 
    IR_r <- get_interest_raroc(
      Purchase_year = year, Fuel = "BET",
      upfront_cost = upfront_cost, subsidy = 0,
      residual_value = max_rvg_s,
      PD = PD_BET_after$PD[PD_BET_after$year == year],
      region = "Europe", company = "industry")

    NPV_r <- get_NPV_fin_instrument(
      Purchase_year = year, subsidy_type = "fixed_percentage",
      subsidy = 0, interest_high = IR_r, interest_low = IR_r,
      residual_value = max_rvg_s, road_tolls_exemp = 0,
      second_hand_value = RV_adj$RV[RV_adj$year == year + 5])

    rv_rvg <- rbind(rv_rvg, data.frame(year = year, rv_d = rv_d,
                                       NPV_gov = NPV_r$residual_value_subsidy))

    # Combined
    IR_c <- get_interest_raroc(
      Purchase_year = year, Fuel = "BET",
      upfront_cost = upfront_cost, subsidy = max_subsidy_combined_s,
      residual_value = max_rvg_combined_s,
      PD = PD_BET_after$PD[PD_BET_after$year == year],
      region = "Europe", company = "industry")

    NPV_c <- get_NPV_fin_instrument(
      Purchase_year = year, subsidy_type = "fixed_percentage",
      subsidy = max_subsidy_combined_s,
      interest_high = IR_c, interest_low = IR_c,
      residual_value = max_rvg_combined_s, road_tolls_exemp = 0,
      second_hand_value = RV_adj$RV[RV_adj$year == year + 5])

    rv_com <- rbind(rv_com, data.frame(year = year, rv_d = rv_d,
                                       NPV_gov = NPV_c$residual_value_subsidy + NPV_c$purchase_subsidy))
  }
}
```

```{r}
#Plotting
incentive_levels <- c("Subsidy (50%)", "Subsidy (30%) + RVG (25%)", "RVG (25%)")

env_RV <- bind_rows(
  rv_sub %>% group_by(year) %>% summarise(NPV_min = min(NPV_gov), NPV_max = max(NPV_gov), .groups = "drop") %>%
    mutate(incentive = factor("Subsidy (50%)", levels = incentive_levels), source = "RV ±5pp"),
  rv_rvg %>% group_by(year) %>% summarise(NPV_min = min(NPV_gov), NPV_max = max(NPV_gov), .groups = "drop") %>%
    mutate(incentive = factor("RVG (25%)", levels = incentive_levels), source = "RV ±5pp"),
  rv_com %>% group_by(year) %>% summarise(NPV_min = min(NPV_gov), NPV_max = max(NPV_gov), .groups = "drop") %>%
    mutate(incentive = factor("Subsidy (30%) + RVG (25%)", levels = incentive_levels), source = "RV ±5pp")
)

 Gov_expenditure_combined_sensitivity <- ggplot() +
  geom_ribbon(data = env_RV,
              aes(x = year, ymin = NPV_min, ymax = NPV_max,
                  fill = incentive, group = interaction(incentive, source),
                  alpha = source)) +

  geom_segment(data = subsidy_results_gg,
               aes(x = year, xend = year_next, y = NPV_gov, yend = NPV_next,
                   color = factor("Subsidy (50%)", levels = incentive_levels),
                   linetype = TCO_parity), size = 1.2) +
  geom_segment(data = combined_results_gg,
               aes(x = year, xend = year_next, y = NPV_gov, yend = NPV_next,
                   color = factor("Subsidy (30%) + RVG (25%)", levels = incentive_levels),
                   linetype = TCO_parity), size = 1.2) +
  geom_segment(data = rvg_results_gg,
               aes(x = year, xend = year_next, y = NPV_gov, yend = NPV_next,
                   color = factor("RVG (25%)", levels = incentive_levels),
                   linetype = TCO_parity), size = 1.2) +

  geom_point(data = TCO_parity_gg,
             aes(x = year, y = NPV_gov, shape = "Onset of TCO parity"),
             color = "black", size = 3.5) +

  scale_color_manual(values = ITFcolours) +
  scale_fill_manual(values = ITFcolours, guide = "none") +
  scale_alpha_manual(values = c("RV ±5pp" = 0.30)) +
  scale_linetype_manual(values = c("Yes" = "solid", "No" = "dotted")) +
  scale_y_continuous(labels = function(x) format(x, big.mark = " "),
                     expand = c(0, 0), limits = c(0, 200000)) +
  scale_x_continuous(expand = c(0, 0), limits = c(2024, 2030.1)) +
  labs(y = "Government expenditure (€/BET)", x = "Year",
       color = "Financial incentive \n(% of initial purchase price)",
       alpha = "Sensitivity range",
       linetype = "TCO parity", shape = "") +
  theme_classic()

Gov_expenditure_combined_sensitivity
```


