---
title: "TCO Modeling for a 4x4 Truck Tractor"
Paper: Financing Road Freight Electrification: Modeling the Total Cost of Ownership of Battery-Electric Trucks under Residual Value and Interest Rate Uncertainties.
Authors: Juan C. Lopez, Andreas Kopf, Matteo Craglia, and Miguel Jaller
---


#Loading packages
```{r}

library(dplyr)
library(reshape)
library(ggplot2)
library(nlme)
library(Hmisc)
library(tidyverse)
library(stats)
library(scales)
library(FinCal)
library(waterfalls)
library(ggpattern)
library(patchwork)
library(svglite)
```

#Basic setup
```{r}
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)
options(tidyverse.quiet = TRUE)

ITFcolours<- c(
  "#1f77b4",  # strong blue
  "#d62728",  # bold red
  "#5fa00a",  # strong green
  "#ff7f0e",  # vivid orange
  "#9867dd",  # medium purple
  "#aaa000",  # lime yellow
  "#081ed3",  # turquoise
  "#390b79",  # deep indigo
  "#af1493",  # deep pink
  "#ff1147",  # tomato red
  "#2A6B1A",  # forest green
  "#ffcc00",  # bright yellow
  "#19998f",  # teal green
  "#e377c2",  # vibrant pink
  "#9f4500"   # orange-red
)

# Working directory: anchored on this Rmd's folder (portable across machines)
get_script_dir <- function() {
  ki <- try(knitr::current_input(dir = TRUE), silent = TRUE)
  if (!inherits(ki, "try-error") && !is.null(ki) && nzchar(ki)) return(dirname(ki))
  if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
    p <- rstudioapi::getSourceEditorContext()$path
    if (nzchar(p)) return(dirname(p))
  }
  getwd()
}
script_dir <- get_script_dir()
```

#Preparing the input data
```{r}
input_dir  <- file.path(script_dir, "Inputs")
output_dir <- script_dir
```

```{r}
#Other inputs from https://theicct.org/wp-content/uploads/2021/11/tco-bets-europe-1-nov21.pdf

#General for NPV:
ownership_time <- 5 #years
discount_rate <- 0.095
interest_rate <- 0.06
upfront_cost <- 0.2 #Variable added for sensitivity analysis
residual_value <- 0.3 #Diesel
subsidy <- 0 #Subsidy

#Fixed costs:
registration_fee<- 800 #[Eur] // Value for France
DT_price <- data.frame( #[Eur]
  year = c(2024, 2030),
  price = c(134100, 137300)) 

BET_price <- read.csv(file.path(input_dir, "BET-Price-IEA_adjusted.csv")) %>% #Source: https://iea.blob.core.windows.net/assets/0aa4762f-c1cb-4495-987a-25945d6de5e8/GlobalEVOutlook2025.pdf adjusted with projections from https://theicct.org/wp-content/uploads/2021/11/tco-bets-europe-1-nov21.pdf
  dplyr::rename(year = Year, price = BET.Retail.Price..Eur.)
  
#Variable costs or parameters:
ownership_tax <- 950 #[Eur/year] // Value for France
road_use_charge <- 0 #[Eur/year] // Value for France
road_tolls <- 0.32 #[Eur/km] // Value for France
maintenance_costs <- data.frame( #[Eur/100km]
  fuel = c("diesel", "BET"),
  Maintenance_costs = c(18.5, 13.24)
)

diesel_consumption <- data.frame( #[l/100km]
  year = c(2024, 2030),
  fuel_consumption = c(27.7, 27.15) #Source: IICT adjusted with IEA trends 
)

energy_consumption <- data.frame( #[kWh/km]
  year = c(2020, 2030),
  energy_consumption = c(1.38, 0.99)
  )

diesel_price <- data.frame( #[Eur/l]
  year= c(2020,2030),
  price = c(0.905, 0.905)
)

electricity_price <- data.frame( #[Eur/kWh]
  year = c(2020, 2030),
  price = c(0.369, 0.127)
)

VKT_year <- read.csv(file.path(input_dir, "VKT_perYear.csv")) %>% #Source: https://theicct.org/wp-content/uploads/2021/11/tco-bets-europe-1-nov21.pdf
  dplyr::rename(service_year = Service.year, VKT = VKT..km.)
```


#Interpolating data between years

```{r}
years_to_interpolate = 2020:2035

DT_price <- approx(x=DT_price$year, y=DT_price$price, xout = 2024:2030, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, price = y)
 
diesel_consumption <- approx(x=diesel_consumption$year, y=diesel_consumption$fuel_consumption, xout = 2024:2035, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, fuel_consumption = y)

energy_consumption <- approx(x=energy_consumption$year, y=energy_consumption$energy_consumption, xout = years_to_interpolate, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, energy_consumption = y)

diesel_price <- approx(x=diesel_price$year, y=diesel_price$price, xout = years_to_interpolate, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, price = y)

electricity_price <- approx(x=electricity_price$year, y=electricity_price$price, xout = years_to_interpolate, method = "linear", rule = 2) %>% 
  as.data.frame() %>%
  dplyr::rename (year = x, price = y)
```

#Checking the Retail Price
```{r}
price_trucks_plot <- ggplot() +
  geom_line(data=DT_price %>%
              filter(year >= 2024), aes(x=year, y=price, color= "Diesel truck"), linewidth =1.2) + 
  geom_line(data=BET_price, aes(x=year, y=price, color = "BET"), linewidth = 1.2) +
  scale_color_manual(values=ITFcolours) +
  scale_y_continuous(#expand = c(0,0),
    labels = comma) +
  scale_x_continuous(expand = c(0,0)) +
  labs(#title="Purchase price of Trucks 2020 - 2030",
    y="Price (€)", 
    x="Year",
    color = "")+
  theme_classic()

price_trucks_plot
```


#Checking the VKT
```{r}
vkt_trucks <- ggplot() +
  geom_line(data=VKT_year, aes(x=service_year, y=VKT, color = "VKT"), linewidth =1.2) +
  scale_color_manual(values=ITFcolours) +
  scale_y_continuous(#expand = c(0,0),
    labels = comma) +
  scale_x_continuous(expand = c(0,0),
                     labels = number_format(accuracy = 1)) +
  labs(#title="VKT of trucks by service year", 
       y="km/year", 
       x="Service year") +
  theme_classic()+
  theme(legend.position = "none")

vkt_trucks
```


#Calculating the IRR and TCO 



```{r}

###Detailed TCO assessment , including resale value and financing


get_TCO_detailed <- function(Purchase_year, interest, Fuel, upfront_cost, residual_value, subsidy_type, subsidy, road_tolls_exemp){

  if (Fuel == "diesel"){
    vehicle_price <- DT_price$price[DT_price$year==Purchase_year]
  }else if (Fuel == "BET"){
    vehicle_price <- BET_price$price[BET_price$year==Purchase_year]
  }

  #SUBSIDY
  if(subsidy_type == "fixed_percentage") {
    subsidy_value <- subsidy*vehicle_price
  } else if (subsidy_type == "percentage_difference") {
    vehicle_price_diesel <- DT_price$price[DT_price$year==Purchase_year]
    subsidy_value <- (vehicle_price - vehicle_price_diesel)*subsidy
  } else if (subsidy_type == "fixed_amount"){
    vehicle_price_diesel <- DT_price$price[DT_price$year==Purchase_year]
    if (subsidy <= (vehicle_price - vehicle_price_diesel)){
      subsidy_value <- subsidy
    } else {
      subsidy_value <- (vehicle_price - vehicle_price_diesel)
    }
  }

  total_loan <- (vehicle_price-subsidy_value)*(1-upfront_cost)
  
  if (interest == 0){
    annual_payment <- total_loan / ownership_time
  } else {
  annual_payment <- pmt(
    r=interest, n=ownership_time, pv=-total_loan, fv=0) 
  }

  cash_flow_detailed <- data.frame()

  for (year_ow in 0:ownership_time) {

    #Checking toll exemption
    if(Purchase_year + year_ow <= road_tolls_exemp){
      toll_exemp <- 0 #Exempted
    }else {
      toll_exemp <- 1 #No_exempted
    }

    if (year_ow == 0) {
      vehicle_purchase <- (vehicle_price-subsidy_value)*upfront_cost
      taxes <- registration_fee
      maintenance <- 0
      tolls <- 0
      fuel_consumption <- 0

      financing_cost <- 0
      
      remaining_balance <- total_loan

    } else {
      VKT <- VKT_year$VKT[VKT_year$service_year == year_ow]
      
      taxes <- ownership_tax + road_use_charge
      maintenance <- maintenance_costs$Maintenance_costs[maintenance_costs$fuel == Fuel]*VKT/100
      tolls <- road_tolls*VKT*toll_exemp
      if (interest == 0) {
        financing_cost <- 0
      } else {
        financing_cost <- remaining_balance * interest
      }
      vehicle_purchase <- annual_payment - financing_cost
      
      remaining_balance <- remaining_balance - vehicle_purchase

      if (Fuel == "diesel") {
        fuel_consumption <- diesel_consumption$fuel_consumption[diesel_consumption$year == Purchase_year] *
          VKT * diesel_price$price[diesel_price$year == (Purchase_year + year_ow - 1)] / 100
      } else if (Fuel == "BET") {
        fuel_consumption <- energy_consumption$energy_consumption[energy_consumption$year == Purchase_year] *
          VKT * electricity_price$price[electricity_price$year == (Purchase_year + year_ow - 1)]
      }

    }

    if (year_ow == ownership_time) {
      vehicle_resale <- vehicle_price*residual_value
    } else {
      vehicle_resale <- 0
    }

    cash_flow_detailed <- rbind(
      cash_flow_detailed,
      data.frame(
        year = year_ow,
        vehicle_purchase = vehicle_purchase,
        taxes = taxes,
        maintenance = maintenance,
        tolls = tolls,
        fuel_consumption = fuel_consumption,

        financing_cost = financing_cost,
        vehicle_resale = vehicle_resale
      )
    )

  }
  
  if (interest == 0) {
  NPV_detailed = data.frame(
    vehicle_purchase = npv(r = discount_rate, cf = cash_flow_detailed$vehicle_purchase),
    taxes = npv(r = discount_rate, cf = cash_flow_detailed$taxes),
    maintenance = npv(r = discount_rate, cf = cash_flow_detailed$maintenance),
    tolls = npv(r = discount_rate, cf = cash_flow_detailed$tolls),
    fuel_consumption = npv(r = discount_rate, cf = cash_flow_detailed$fuel_consumption),
    
    financing_cost = 0, 
    vehicle_resale = -npv(r = discount_rate, cf = cash_flow_detailed$vehicle_resale)
  ) %>%
    mutate(NPV = vehicle_purchase +
             taxes +
             maintenance +
             tolls +
             fuel_consumption +
             financing_cost +
             vehicle_resale)
  } else {
    NPV_detailed = data.frame(
      vehicle_purchase = npv(r = discount_rate, cf = cash_flow_detailed$vehicle_purchase),
      taxes = npv(r = discount_rate, cf = cash_flow_detailed$taxes),
      maintenance = npv(r = discount_rate, cf = cash_flow_detailed$maintenance),
      tolls = npv(r = discount_rate, cf = cash_flow_detailed$tolls),
      fuel_consumption = npv(r = discount_rate, cf = cash_flow_detailed$fuel_consumption),
      
      financing_cost = npv(r = discount_rate, cf = cash_flow_detailed$financing_cost),
      vehicle_resale = -npv(r = discount_rate, cf = cash_flow_detailed$vehicle_resale)
    ) %>%
      mutate(NPV = vehicle_purchase +
               taxes +
               maintenance +
               tolls +
               fuel_consumption + 
               financing_cost +
               vehicle_resale)
  }
  

  return(NPV_detailed)

}
```


#Calculating the TCO - Diesel
```{r}
diesel_tco <- data.frame()
BET_tco <- data.frame()

tco_BET_diesel <- data.frame()

for (year_purchase in 2024:2030){
  
  result_tco_diesel <- get_TCO_detailed(Purchase_year = year_purchase, Fuel = "diesel", interest= interest_rate, upfront_cost = upfront_cost, residual_value = residual_value, subsidy_type = "fixed_percentage", subsidy = subsidy, road_tolls_exemp= 0)

  result_tco_BET <- get_TCO_detailed(Purchase_year = year_purchase, Fuel = "BET", interest= interest_rate, upfront_cost = upfront_cost, residual_value = 0, subsidy_type = "fixed_percentage", subsidy = subsidy, road_tolls_exemp= 0)
  
  tco_BET_diesel <- rbind(tco_BET_diesel, data.frame(year = year_purchase, (result_tco_BET/result_tco_diesel$NPV)))
  
  diesel_tco <- rbind(diesel_tco, data.frame(year = year_purchase, result_tco_diesel))

  BET_tco <- rbind(BET_tco, data.frame(year = year_purchase, result_tco_BET))
  
  }
```


#Plotting the TCO
```{r}
ggplot() +
  geom_line(data=diesel_tco, aes(x=year, y=NPV, color= "Diesel Truck"), size =1.2) + 
  geom_line(data=BET_tco, aes(x=year, y=NPV, color = "Battery Electric Truck"), size = 1.2) +
  scale_color_manual(values=ITFcolours) +
  scale_y_continuous(#expand = c(0,0),
                     labels = comma) +
  scale_x_continuous(expand = c(0,0),
                     labels = number_format(accuracy = 1)) +
  labs(title="TCO of Trucks 2020 - 2030", y="Price [Eur]", 
       x="Year",
       color = "")+
  theme_classic()
```



## Plot per component
```{r}
tco_total <- bind_rows(diesel_tco %>% mutate(truck_type = "Diesel"), BET_tco %>% mutate(truck_type = "BET"))

tco_long <- tco_total %>%
  select(-NPV) %>%  # Remove NPV since it's the total
  pivot_longer(cols = -c(year, truck_type), names_to = "TCO_component", values_to = "cost") %>%
  mutate(truck_type = case_when(truck_type == "BET" ~ "BET",
                                .default = truck_type))
```

```{r}
tco_long$TCO_component <- recode(tco_long$TCO_component, #Renaming the components for plotting
                                 "financing_cost" = "Financing (IR=6%)",
                                 "vehicle_purchase" = "Vehicle purchasing",
                                 "taxes" = "Taxes",
                                 "maintenance" = "Maintenance",
                                 "fuel_consumption" = "Fuel",
                                 "tolls" = "Tolls",
                                 "vehicle_resale" = "Residual Value" 
)


# Plot stacked bar chart
ggplot(tco_long, aes(x = year, y = cost, fill = TCO_component)) +
  geom_bar(stat = "identity") +
  facet_wrap(~truck_type) +
  scale_fill_manual(values=ITFcolours) +
  scale_y_continuous(expand = c(0,0),
                     labels = scales::comma) +

  labs(
    title = "TCO of Trucks 2024 - 2030 per component",
    x = "Year",
    y = "TCO [EUR]",
    fill = "TCO Component"
  ) +
  theme_classic() #+
```

```{r}
## Recalculating the TCO with an IR of 15% to update the previous plot
BET_tco_IR15 <- data.frame()

for (year_purchase in 2024:2030){
  result_tco_BET_IR15 <- get_TCO_detailed(Purchase_year = year_purchase, Fuel = "BET", interest= 0.15, upfront_cost = upfront_cost, residual_value = 0, subsidy_type = "fixed_percentage", subsidy = subsidy, road_tolls_exemp= 0)
  
  BET_tco_IR15 <- rbind(BET_tco_IR15, data.frame(year = year_purchase, result_tco_BET_IR15))
}

BET_tco_complete <- BET_tco %>%
  left_join(BET_tco_IR15 %>% 
              dplyr::select(c(year, financing_cost)) %>%
              dplyr::rename("financing_cost_15" = "financing_cost"),
            by = "year") %>%
  mutate(financing_cost_15 = financing_cost_15-financing_cost) 

```

```{r}
## Replotting:

tco_total_IR15 <- bind_rows(diesel_tco %>% mutate(truck_type = "Diesel"), BET_tco_complete %>% mutate(truck_type = "BET"))

tco_long_IR15 <- tco_total_IR15 %>%
  select(-NPV) %>%  # Remove NPV since it's the total
  pivot_longer(cols = -c(year, truck_type), names_to = "TCO_component", values_to = "cost") %>%
  mutate(truck_type = case_when(truck_type == "BET" ~ "BET",
                                .default = truck_type))
```

```{r}
tco_long_IR15$TCO_component <- recode(tco_long_IR15$TCO_component, #Renaming the components for plotting
                                 "financing_cost_15"="Financing (IR=15%)",
                                 "financing_cost" = "Financing (IR=6%)",
                                 "vehicle_purchase" = "Vehicle purchasing",
                                 "taxes" = "Taxes",
                                 "maintenance" = "Maintenance",
                                 "fuel_consumption" = "Fuel",
                                 "tolls" = "Tolls",
                                 "vehicle_resale" = "Residual Value" 
)


# Plot stacked bar chart
tco_comparison <- ggplot(tco_long_IR15, aes(x = year, y = cost, fill = TCO_component)) +
  geom_bar_pattern(stat = "identity",
    aes(pattern = TCO_component),
    
    # Visual settings for the pattern (the diagonal lines)
    pattern_color = "black", 
    pattern_fill = "black",       
    pattern_angle = 45,  
    pattern_density = 0.1,  
    pattern_spacing = 0.02,       
    pattern_alpha = 0.3  
  ) +
  facet_wrap(~truck_type) +
  scale_fill_manual(values = c(
    "Financing (IR=6%)" = scales::alpha(ITFcolours[5], 0.95),
    "Financing (IR=15%)" = scales::alpha(ITFcolours[5], 0.3),
    "Vehicle purchasing" = scales::alpha(ITFcolours[2], 0.95),
    "Taxes" = scales::alpha(ITFcolours[1], 0.95),
    "Maintenance" = scales::alpha(ITFcolours[4], 0.95),
    "Fuel" = scales::alpha(ITFcolours[3], 0.95),
    "Tolls" = scales::alpha(ITFcolours[6], 0.95),
    "Residual Value" = scales::alpha(ITFcolours[7], 0.95)
  )) +
  scale_pattern_manual(values = c(
    "Financing (IR=15%)" = "stripe",  # This gets the texture
    "Financing (IR=6%)" = "none",     # These get no texture
    "Vehicle purchasing" = "none",
    "Taxes" = "none",
    "Maintenance" = "none",
    "Fuel" = "none",
    "Tolls" = "none",
    "Residual Value" = "none"
  )) +
  
  scale_y_continuous(expand = c(0,0),
                     labels = function(x) format(x, big.mark = " ", scientific = FALSE))+

  labs(
    #title = "TCO of Trucks 2024 - 2030 per component",
    x = "Year of purchase",
    y = "TCO (€)",
    fill = "TCO Component",
    pattern = "TCO Component"
  ) +
  theme_classic() #+

tco_comparison
```


## BET/diesel TCO

```{r}
ggplot(tco_BET_diesel, aes(x = year, y = NPV)) +
  geom_line(aes(color = "BET/Diesel Ratio"), size = 1.2) + 
  geom_hline(yintercept = 1, linetype = "dashed", color = "red", size = 1) +
  scale_color_manual(values=ITFcolours) +
  scale_y_continuous(#expand = c(0,0),
                     labels = comma) +
  scale_x_continuous(expand = c(0,0),
                     labels = number_format(accuracy = 1)) +
  labs(title="Ratio of BET/Diesel TCO 2024-2030", y="TCO ratio BET/Diesel [-]", 
       x="Year",
       color = "")+
  theme_classic()
```



## Waterfall diagram

```{r}
tco_comparison_long <- data.frame(
  year = BET_tco$year,
  TCO_diesel = diesel_tco$NPV,
  BET_tco[!names(BET_tco) %in% c("year", "NPV")] - diesel_tco[!names(diesel_tco) %in% c("year", "NPV")]) %>% 
  mutate(across(-year, ~round(.x/TCO_diesel*100, 1))) %>%
  filter(year == 2025) %>%
  pivot_longer(cols = -c("year"), names_to = "TCO_component", values_to = "cost") %>%
  select(-"year")

tco_comparison_long$TCO_component <- recode(tco_comparison_long$TCO_component, #RENAMING FOR PLOTTING
                                            "TCO_diesel" = "TCO Diesel",
                                            "financing_cost" = "Financing",
                                            "vehicle_purchase" = "Vehicle purchase",
                                            "taxes" = "Taxes",
                                            "maintenance" = "Maintenance",
                                            "fuel_consumption" = "Fuel",
                                            "tolls" = "Tolls",
                                            "vehicle_resale" = "Residual Value" 
)

#Reordering the components for the plot
tco_comparison_long <- tco_comparison_long %>%
  slice(c(
    which(TCO_component == "TCO Diesel"),
    which(TCO_component == "Vehicle purchase"),
    which(TCO_component == "Taxes"),
    which(TCO_component == "Maintenance"),
    which(TCO_component == "Tolls"),
    which(TCO_component == "Fuel"),
    which(TCO_component == "Residual Value"), 
    which(TCO_component == "Financing")
  ))

 
waterfall_TCO_1 <- waterfall(tco_comparison_long, calc_total = TRUE, total_axis_text = "TCO BET", rect_text_size = 1, fill_by_sign = FALSE, fill_colours = alpha(ITFcolours, 0.8)) + 
  scale_y_continuous(expand = c(0,0),
                     labels = comma,
                     limits = c(0,150)) +
  labs(title="TCO comparison BETs vs Diesel 2025", y="TCO ratio BET/Diesel (%)", 
       x="TCO component")+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

waterfall_TCO_1
```

##Waterfall with an IR of 15%
```{r}
tco_comparison_long_IR15 <- data.frame(
  year = BET_tco_IR15$year,
  TCO_diesel = diesel_tco$NPV,
  BET_tco_IR15[!names(BET_tco_IR15) %in% c("year", "NPV")] - diesel_tco[!names(diesel_tco) %in% c("year", "NPV")]) %>% 
  mutate(across(-year, ~round(.x/TCO_diesel*100, 1))) %>%
  filter(year == 2025) %>%
  pivot_longer(cols = -c("year"), names_to = "TCO_component", values_to = "cost") %>%
  select(-"year")

tco_comparison_long_IR15$TCO_component <- recode(tco_comparison_long_IR15$TCO_component, #RENAMING FOR PLOTTING
                                            "TCO_diesel" = "TCO Diesel",
                                            "financing_cost" = "Financing",
                                            "vehicle_purchase" = "Vehicle purchase",
                                            "taxes" = "Taxes",
                                            "maintenance" = "Maintenance",
                                            "fuel_consumption" = "Fuel",
                                            "tolls" = "Tolls",
                                            "vehicle_resale" = "Residual Value" 
)

#Reordering the components for the plot
tco_comparison_long_IR15 <- tco_comparison_long_IR15 %>%
  slice(c(
    which(TCO_component == "TCO Diesel"),
    which(TCO_component == "Vehicle purchase"),
    which(TCO_component == "Taxes"),
    which(TCO_component == "Maintenance"),
    which(TCO_component == "Tolls"),
    which(TCO_component == "Fuel"),
    which(TCO_component == "Residual Value"),
    which(TCO_component == "Financing") 
  ))

 
waterfall_TCO_2 <- waterfall(tco_comparison_long_IR15, calc_total = TRUE, total_axis_text = "TCO BET", rect_text_size = 1, fill_by_sign = FALSE, fill_colours = alpha(ITFcolours, 0.8)) + 
  scale_y_continuous(expand = c(0,0),
                     labels = comma,
                     limits = c(0,140)) +
  labs(title="TCO comparison BETs vs Diesel 2025", y="TCO ratio BET/Diesel (%)", 
       x="TCO component")+
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

waterfall_TCO_2
```


```{r}
#Calculating the differences:
TCO1_max <- max(waterfall_TCO_1[["data"]] %>% filter(x == "TCO BET") %>% pull(y))
TCO2_max <- max(waterfall_TCO_2[["data"]] %>% filter(x == "TCO BET") %>% pull(y))

#Adding them to the plot
TCO_waterfall <- waterfall_TCO_1 + 
  geom_rect_pattern(
    aes(xmin = 7.65, 
        xmax = 8.35, 
        ymin = TCO1_max, 
        ymax = TCO2_max,  
        pattern = "Effects of increasing the IR from 6% to 15%"),
    fill = alpha(ITFcolours[8], 0.02),
    color = "black",  
    linewidth = 0.3,

    pattern_color = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.005,
    pattern_spacing = 0.03,
    pattern_alpha = 0.05
  ) +
 
  geom_rect_pattern(
    aes(xmin = 8.65, 
        xmax = 9.35, 
        ymin = TCO1_max, 
        ymax = TCO2_max,
        pattern = "Effects of increasing the IR from 6% to 15%"),
    fill = alpha("black", 0.02),
    color = "black",  
    linewidth = 0.3,
    pattern_color = "black",
    pattern_fill = "black",
    pattern_angle = 45,
    pattern_density = 0.005,
    pattern_spacing = 0.03,
    pattern_alpha = 0.05
  ) + 
  
  scale_pattern_manual(
    values = c("Effects of increasing the IR from 6% to 15%" = "stripe")
  ) +
  
  guides(
    pattern = guide_legend(
      override.aes = list(
        pattern_alpha = 0.3,  
        pattern_density = 0.01,  
        pattern_spacing = 0.02, 
        fill = NA
      )
    )
  ) +
  
  geom_segment(
    aes(x = 8.35, 
        xend = 8.65,
        y = TCO2_max, 
        yend = TCO2_max),
    color = "black",
    linetype = "dashed",
    linewidth = 0.3,
  ) +
  annotate("text", 
           x = 8, 
           y = (TCO1_max + TCO2_max) / 2, 
           label = round(TCO2_max - TCO1_max, 1), 
           size = 3,
           fontface = "bold", 
           color = "black") +
  
  annotate("text", 
           x = 9, 
           y = (TCO1_max + TCO2_max) / 2, 
           label = round(TCO2_max - TCO1_max, 1), 
           size = 3,
           fontface = "bold", 
           color = "black") +
  
  labs(title=NULL,
       pattern = NULL) +  # Removes the legend title (keeps only the label)
  theme(legend.position = "bottom")
  

TCO_waterfall
```


#SENSITIVITY ANALYSIS:

## 1. Residual value guarantee
```{r}
BET_tco_sensitivity <- data.frame()

for (residual_value_BET in unique(sort(c(seq(0, 0.8, by = 0.1), 0.4)))) { # I am adding the sequence from 0.12 to 0.96 and 0.4 for visualization purposes
  for (year_purchase in 2024:2030){
    result_tco_BET_sensitivity <- get_TCO_detailed(Purchase_year = year_purchase, Fuel = "BET", interest= interest_rate, upfront_cost = upfront_cost, residual_value = residual_value_BET, subsidy_type = "fixed_percentage", subsidy = subsidy, road_tolls_exemp= 0)
    
    diesel_tco_sensitivity <- diesel_tco %>%
      filter(year == year_purchase) %>%
      select(-year)
    
    BET_tco_sensitivity <- rbind(BET_tco_sensitivity, data.frame(year = year_purchase, (result_tco_BET_sensitivity/diesel_tco_sensitivity), residual_value = residual_value_BET))
  }
}
```

```{r}
#Preparing data for visualization

plot1 <- ggplot() +
  
  geom_line(data = BET_tco_sensitivity %>%
              filter(residual_value %in% c(0, 0.1, 0.2, "0.3", 0.4)),
            aes(x = year, y = NPV,
                color = factor(residual_value*100, levels = c("0", "10", "20", "30", "40"))),
            linewidth  = 1.2) +
  
  geom_hline(yintercept = 1, color = "red", linetype = "dashed", size = 1) +
  
  annotate("text", x = 2025, y = 0.95, label = "Diesel TCO", color = "red", size = 4) +
  
  scale_color_manual(
    name = "RV (%)",
    values = ITFcolours) +#c("0" = 0.1, "20" = 0.4 , "40" = 0.7, "60" = 1)) +
  
  scale_y_continuous(expand = c(0,0),
                     labels = scales::number_format(accuracy = 0.1),
                     limits = c(0.8, 1.65),
                     breaks = seq(0.8, 1.65, 0.2)) +
  scale_x_continuous(expand = c(0,0))+#,

  labs(#title="TCO of Trucks on a 5-year horizon", y="TCO ratio BET/Diesel (-)", 
    title = NULL,   
    x="Year of purchase",
    y = "TCO ratio BET/Diesel (-)",
    color = "",
    fill = "BET's Residual Value (%)") +
  theme_classic() +
  theme(legend.position = "bottom")

plot1

```


## 2. Interest rate
```{r}
BET_tco_sensitivity <- data.frame()

for (interest_BET in c(0.01,seq(0, 0.30, by = 0.05))) {
  for (year_purchase in 2024:2030){
    
    result_tco_BET_sensitivity <- get_TCO_detailed(Purchase_year = year_purchase, Fuel = "BET", interest= interest_BET, upfront_cost = upfront_cost, residual_value = 0, subsidy_type = "fixed_percentage", subsidy = subsidy, road_tolls_exemp= 0)
    
    diesel_tco_sensitivity <- diesel_tco %>%
      filter(year == year_purchase) %>%
      select(-year)
    
    BET_tco_sensitivity <- rbind(BET_tco_sensitivity, data.frame(year = year_purchase, (result_tco_BET_sensitivity/diesel_tco_sensitivity), interest = interest_BET))
  }
}
```

#Plotting sensitivity

```{r}
#Preparing data for visualization

plot2 <- ggplot() +
  geom_line(data = BET_tco_sensitivity %>%
              filter(interest %in% c(0, 0.05, 0.1, "0.15", 0.2)),
            aes(x = year, y = NPV,
                color = factor(interest * 100, 
                               levels = rev(c("0", "5", "10", "15", "20")))),  # reversed
            linewidth = 1.2) +
  
  geom_hline(yintercept = 1, color = "red", linetype = "dashed", size = 1) +
  annotate("text", x = 2025, y = 1.05, label = "Diesel TCO", color = "red", size = 4) +
  
  scale_color_manual(
    name = "IR (%)",
    values = ITFcolours # 4 interest levels = 4 colors from position 5 onward
  ) +
  
  scale_y_continuous(expand = c(0, 0),
                     labels = scales::number_format(accuracy = 0.1),
                     limits = c(0.8, 1.65),
                     breaks = seq(0.8, 1.65, 0.2)) +
  scale_x_continuous(expand = c(0, 0)) +
  
  labs(#title = "TCO of Trucks on a 5-year horizon", 
    title = NULL,
    y = "TCO ratio BET/Diesel (-)", 
    x = "Year of purchase",
    color = "") +
  
  theme_classic() + 
  theme(legend.position = "bottom") +
  guides(color = guide_legend(reverse = TRUE))

plot2

```


```{r}
## Putting the plots together:

TCO_RV_vs_IR <- plot1 + plot2 +
  plot_layout(guides = "collect") & 
  theme(legend.position = 'bottom')

TCO_RV_vs_IR
```


##--------------- RAROC AND RoE ------------

```{r}
raroc_data_region <- read.csv(file.path(input_dir, "PD_RR_Regions.csv")) #Source: https://www.eib.org/attachments/lucalli/20240217_default_and_recovery_statistics_private_and_public_lending_1994_2023_en.pdf#page=9.08
```


#FUNCTION TO CALCULATE THE INTEREST NEEDED TO REACH CERTAIN LEVEL OF RETURN ON EQUITY

```{r}
## Followed methodology: https://roopya.money/introduction-to-loan-pricing-and-raroc-risk-adjusted-return-on-capital/
###OUTPUT: INTEREST RATE
get_interest_raroc <- function(Purchase_year, Fuel, upfront_cost, subsidy, residual_value, prob_default, region){
  
  if (Fuel == "Diesel") {
    vehicle_price <- DT_price$price[DT_price$year == Purchase_year]
    # PD <- 0.045 #Source: https://www.eib.org/attachments/lucalli/20240217_default_and_recovery_statistics_private_and_public_lending_1994_2023_en.pdf#page=9.08
    PD <- raroc_data_region$PD[raroc_data_region$Region == region]
    # RecoveryRate <- 0.672 #Source: https://www.eib.org/attachments/lucalli/20240217_default_and_recovery_statistics_private_and_public_lending_1994_2023_en.pdf#page=9.08
    RecoveryRate <- raroc_data_region$recovery_rate[raroc_data_region$Region == region]
  } else if (Fuel == "BET") {
    vehicle_price <- BET_price$price[BET_price$year == Purchase_year]
    PD <- prob_default
    RecoveryRate <- residual_value
  }
  
  total_loan <- (1 - upfront_cost - subsidy) * vehicle_price
  
  LGD <- 1 - RecoveryRate #Source: https://www.investopedia.com/terms/l/lossgivendefault.asp
  EL <- total_loan*PD*LGD #Expected loss
  
  Cost_capital <- 0.0357 #Reflects bank's funding costs. For now Aprox to Return on Equity. Source: https://www.ecb.europa.eu/pub/pdf/scpops/ecb.op254~664ed99e11.en.pdf
  Economic_capital_requirement <- 0.08 #The source indicates it must not be lower than 8%. Source: https://www.bis.org/publ/bcbs128b.pdf 
  Economic_capital = Economic_capital_requirement*total_loan
  
  raroc <- 0.10
  
  Op_expenses <- 0.689 #As a % of net interest income. Source: https://databank.worldbank.org/source/global-financial-development/Series/GFDD.EI.07

  NII <- (raroc*Economic_capital + EL)/(1-Op_expenses)
  
  interest_rate_output <- (NII + Cost_capital*Economic_capital)/total_loan
  
  return(interest_rate_output)
}
```

#CALCULATING ACROSS DIFFERENT PD AND RVG
```{r}
interest_BET_sensitivity <- data.frame()
interest_diesel <-data.frame()

for (region in raroc_data_region$Region) {
  interest_diesel_result <- get_interest_raroc(Purchase_year = 2030, Fuel="Diesel", upfront_cost = upfront_cost, subsidy = subsidy, residual_value = residual_value, prob_default = PD, region = region)
  
  interest_diesel <- rbind(interest_diesel, data.frame(interest = interest_diesel_result, region = region ))
}
  
for(residual_value_BET in seq(0, 0.7, by = 0.1)) {
  for(PD_BET in c(0.02, 0.05, 0.1, 0.15, 0.2)){
    interest_BET_result <- get_interest_raroc(Purchase_year = 2030, Fuel="BET", upfront_cost = upfront_cost, subsidy = subsidy, residual_value = residual_value_BET, prob_default = PD_BET, region = "Sector_Avg")
    
    interest_BET_sensitivity <- rbind(interest_BET_sensitivity, data.frame(interest = interest_BET_result, Truck = "BET", prob_default = PD_BET, residual_value = as.numeric(residual_value_BET)))
  }
}

```

## Plotting the results
```{r}

RAROCvsPDvsRV <- ggplot() +
  geom_bar(data =interest_BET_sensitivity %>%
             filter(residual_value %in% c(0, 0.1, 0.2, "0.3", 0.4)),
           aes(x = factor(prob_default*100), 
               y = interest*100, 
               fill = factor(residual_value*100)), 
           stat = "identity", 
           position = "dodge", 
           alpha = 0.9) +
  geom_hline(yintercept = interest_diesel$interest[interest_diesel$region=="Sector_Avg"]*100, linetype = "dashed", color = "red", size = 1) +
  annotate("text", x = 1.2, y = (interest_diesel$interest[interest_diesel$region=="Sector_Avg"] + 0.025)*100, label = "IR = 6%", color = "red", size = 4) +
  scale_color_manual(values = ITFcolours) +
  scale_fill_manual(values = ITFcolours) +
  scale_y_continuous(expand = c(0,0))+
  # labels = percent_format(accuracy = 1)) +
  labs(#title = "Interest Rate required to reach a RoE of 10%",
    title = NULL,
    x = "Probability of default (%)", y = "Interest Rate (%)",
    fill = "Residual Value (%)") +
  theme_classic()

RAROCvsPDvsRV 
```
